Exploring dimensions of biodiversity in Japanese ferns

### 2021-08-19 ms/manuscript.Rmd
Update MS
Fix description of some results

**Results** We observe three major patterns in the spatial distribution of biodiversity: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity and phylogenetic endemism, which are highest in small southern islands, and that of relative phylogenetic and functional diversity, which are high at both high and low latitudes, and low at mid-latitudes. 
Grid-cells were grouped in to three (phylogenetic) or four (taxonomic) major bioregions. 
Temperature and apomixis were identified as likely drivers of patterns of taxonomic and phylogenetic diversity.
Conservation status (% protected area) is higher than the overall rate for grid-cells with significantly high biodiversity.

@@

However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, and such datasets are rare on the regional (country) scale. 
The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled. 
One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].

@@

We used traits originally compiled for identification of ferns and lycophytes of Japan following Ebihara [-@Ebihara2016b; -@Ebihara2017]. 
Continuous traits were measured on 10 randomly selected specimens per species. 
Fertile (`r ie`, spore-bearing) and sterile fronds were measured separately, then averaged. 
Qualitative traits were scored by observing voucher specimens. 
All qualitative traits were scored in binary format; `r eg`, a single trait with three states was formatted as three binary traits. 
From the original trait list, we selected only traits with putative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.

@@

We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`). 
We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
This approach reflects the idea that the amount of area with a given climate regime enclosing each grid cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
The R package `r pack("sf")` was used for GIS analyses.

@@

All measures of biodiversity and CANAPE were carried out using the `r pack("canaper")` package [@Nitta2021], which calculates raw biodiversity metrics using functions in the `r pack("phyloregion")` package [@Daru2020a] and randomizations using the `r pack("picante")` package [@Kembel2010].

@@

Grid-cells with significantly low PD were almost all located on Honshu, the main island of Japan (`r figure("rand-diversity")`a). 
A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Izu Peninsula, Kii Peninsula). 
The area near Yakushima includes cells with both significantly high and low PD. 
Grid-cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu (`r figure("rand-diversity")`b). 
Grid-cells with significantly high RPD are located both in the southern islands and northern Honshu and Hokkaido. 
Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity, but with more significant cells for FD than PD and fewer significant cells for RFD than RPD (`r figure("rand-diversity")`c, d).

@@

All models were significant relative to null models without any fixed effects `r s_table("lrt-null")`. 
The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`a--e), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`f--g); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models. 
Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`a). 
Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`b--e). 
SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`g); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`f). 
Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD (`r s_table("aic")`).

@@

The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`). 
Areas with high PE were the best protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total). 
All cells with significantly high biodiversity had a greater percentage of area protected than the overall rate for Japan (`r scales::percent(total_percent_protected_high, accuracy = 0.1)` high, `r scales::percent(total_percent_protected_medium, accuracy = 0.1)` medium, `r scales::percent(total_percent_protected_all, accuracy = 0.1)` total).

@@

Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu (`r figure("rand-diversity")`)b. 
The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`a). The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`b) and *Dryopteris* (`r s_figure("genus-map")`d).

@@

## Author Contribution

@@

# Make map
japan_map_plot <- ggplot(japan) + 
  geom_sf(size = 0.25) +
  geom_segment(data = japan_lines, aes(x = lon, y = lat, xend = lon_end, yend = lat_end, linetype = query)) +
  scale_linetype_manual(values = c("Tsugaru Strait" = "solid", "Tokara Strait" = "longdash")) +
  geom_richtext(
    data = japan_points, 
    aes(x = lon_adj, y = lat_adj, label = label),
    size = 2.5,
    fill = NA, label.color = NA, # remove background and outline
    label.padding = grid::unit(rep(0, 4), "pt") # remove padding)
  ) +
  geom_point(
    data = filter(japan_points, query == "Tokyo"),
    aes(x = lon, y = lat)
  ) +
  map_theme() +
  theme(legend.position = "none")

ggsave("plot.pdf", japan_map_plot, width = 5, height = 5, units = "in")
```

@@

**`r figure("div-by-tax-region")`**. Phylogenetic and morphological diversity of the ferns of Japan by bioregion. (a), Standard effect size (SES) of phylogenetic diversity (PD). (b), SES of relative PD (RPD). (c), SES of functional diversity (FD). (d), SES of relative FD (RFD). (e), *p*-score of observed phylogenetic endemism (PE) relative to 999 random communities. Only top four clusters with the most sites shown.

@@

**`r figure("conserv-status")`**. Percent of land area with protected status for grid cells with significantly high biodiversity. Biodiversity metrics include phylogenetic diversity (PD), functional diversity (FD), relative PD (RPD), relative FD (RFD), and phylogenetic endemism (PE). Significance of biodiversity metrics assessed by a one-tailed test comparing observed values to a null distribution of 999 random values (see Methods). Vertical lines indicate total percent area protected across Japan regardless of degree of biodiversity; dashed, high protection level (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`); dotted, medium protection level (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`).

>>>>

**Results** We observe three major patterns in the spatial distribution of biodiversity: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity and phylogenetic endemism, which are highest in small southern islands, and that of relative phylogenetic and functional diversity, which are high at both high and low latitudes, and low at mid-latitudes. 
Grid-cells were grouped in to three (phylogenetic) or four (taxonomic) major bioregions. 
Temperature and apomixis were identified as likely drivers of patterns of taxonomic and phylogenetic diversity.
Conservation status (% protected area) is generally higher than the overall rate for grid-cells with significantly high biodiversity.

@@

However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, and such datasets are rare on the regional (country) scale. 
The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b]. 
One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].

@@

We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017]. 
Continuous traits were measured on 10 randomly selected specimens per species. 
Fertile (`r ie`, spore-bearing) and sterile fronds were measured separately, then averaged. 
Qualitative traits were scored by observing voucher specimens. 
All qualitative traits were scored in binary format; `r eg`, a single trait with three states was formatted as three binary traits. 
From the original trait list, we selected only traits with putative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.

@@

We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`). 
We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
This approach reflects the idea that the amount of area with a given climate regime enclosing each grid cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
The R package `r pack("sf")` was used for GIS analyses [@Pebesma2018].

@@

All measures of biodiversity and CANAPE were carried out using the `r pack("canaper")` package [@Nitta2021a], which calculates raw biodiversity metrics using functions in the `r pack("phyloregion")` package [@Daru2020a] and randomizations using the `r pack("picante")` package [@Kembel2010].

@@

Grid-cells with significantly low PD were almost all located on Honshu, the main island of Japan (`r figure("rand-diversity")`a). 
A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Kii Peninsula). 
Grid-cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu (`r figure("rand-diversity")`b). 
Grid-cells with significantly high RPD are located both in the southern islands and northern Honshu and Hokkaido. 
Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity (`r figure("rand-diversity")`c, d).

@@

All models were significant relative to null models without any fixed effects (`r s_table("lrt-null")`). 
The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`a--e), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`f--g); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models. 
Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`a). 
Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`b--e). 
SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`g); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`f). 
Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD and SES of PD (`r s_table("aic")`).

@@

The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`). 
Areas with high PE were the best protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total). 
Cells with significantly high biodiversity generally had a greater percentage of area protected than the overall rate for Japan (`r scales::percent(total_percent_protected_high, accuracy = 0.1)` high, `r scales::percent(total_percent_protected_medium, accuracy = 0.1)` medium, `r scales::percent(total_percent_protected_all, accuracy = 0.1)` total); only FD was lower than the overall rate for highly protected areas.

@@

Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu (`r figure("rand-diversity")`b). 
The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`a). The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`b) and *Dryopteris* (`r s_figure("genus-map")`d).

@@

@@

# Make map
japan_map_plot <- ggplot(japan) + 
  geom_sf(size = 0.25) +
  geom_segment(data = japan_lines, aes(x = lon, y = lat, xend = lon_end, yend = lat_end, linetype = query)) +
  scale_linetype_manual(values = c("Tsugaru Strait" = "solid", "Tokara Strait" = "longdash")) +
  geom_richtext(
    data = japan_points, 
    aes(x = lon_adj, y = lat_adj, label = label),
    size = 2.5,
    fill = NA, label.color = NA, # remove background and outline
    label.padding = grid::unit(rep(0, 4), "pt") # remove padding)
  ) +
  geom_point(
    data = filter(japan_points, query == "Tokyo"),
    aes(x = lon, y = lat)
  ) +
  map_theme() +
  theme(legend.position = "none")
```

@@

**`r figure("div-by-tax-region")`**. Phylogenetic and morphological diversity of the ferns of Japan by taxonomic bioregion. (a), Standard effect size (SES) of phylogenetic diversity (PD). (b), SES of relative PD (RPD). (c), SES of functional diversity (FD). (d), SES of relative FD (RFD). (e), *p*-score of observed phylogenetic endemism (PE) relative to 999 random communities. Only top four clusters with the most sites shown.

@@

**`r figure("conserv-status")`**. Percent of land area with protected status for grid cells with significantly high biodiversity. Biodiversity metrics include phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE). Significance of biodiversity metrics assessed by a one-tailed test comparing observed values to a null distribution of 999 random values (see Methods). Vertical lines indicate total percent area protected across Japan regardless of degree of biodiversity; dashed, high protection level (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`); dotted, medium protection level (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`).
